Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-18 paper/uist_2016/paper.tex
dumped numbers in memorability section; fixed typos in the table

			\multicolumn{1}{|c|}{}                               & \multicolumn{1}{c|}{}                        & \multicolumn{1}{c|}{mean}  & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}66.67\%}      & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}42.02\%}     & \multicolumn{1}{c|}{50.83\%}                              & \multicolumn{1}{c|}{47.46\%}                             & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}100.00\%}          & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}55.00\%}        & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}100.00\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}75.00\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}72.22\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}43.66\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}51.56\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}10.00\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}27.66\%}      & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}64.20\%}     \\ \cline{3-17} 

@@

\subsection{Memorability}
Rewinds are meant to engage a person's memory. To find out more about what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo or recall the trip portrayed by a Rewind. Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the variables of distance from the participant's home, visit frequency, presence of travel companions, regular GPS usage, and whether the participant considered themselves a frequent traveler.

\subsubsection{Rewinds versus Photos}
When participants were shown a street side photo from their location history, they were asked whether they recognized the location in the photo. In 50\% of cases, participants said they did. When they were shown the Rewind associated with that photo, nearly every case had recognizable locations---97\%. This result is expected since Rewinds contain more location information than static photos, so if participants do not recognize some depicted locations, they still have many other chances to pick out locations they remember. Rewind memorability as calculated by whether participants could identify the particular trip they were on from watching the Rewind was 54\%. However, there were cases where the participant would only remember the photo after seeing its accompanying Rewind, which provided a context. More than one participant exclaimed after watching a Rewind, ``I recognize the photo now!''

\subsubsection{Distance}
The distance away from the participant’s home had differing effects on photo memorability and Rewind memorability (Figure~\ref{fig:distancememory}). Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from the participant's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from the participant's home location.

Photo memorability and distance had a clear linear relationship. As the distance increased, photo memorability decreased, suggesting that locations further away from home were less likely to be remembered. Remembrance was 100\% for Distance Group 0, or home. One participant recognized a photo because her ``husband's car is in the driveway.'' Remembrance dropped to 74\% for Distance Group 1, then down to 10\% for Distance Group 5, which represented the locations furthest away from home. These percentages are statistically significant, with a p-value of 0.0028.

@@

\subsubsection{Visit Frequency}
Location memorability and trip memorability are affected differently by frequency of visits (Figure~\ref{fig:frequencymemory}). Location memorability is highest for places that are most frequently visited by the participant, while trip memorability is highest for places that are least frequently visited. The more frequently a place was visited, the more likely it was that the participant remembered the location when it was shown to them at the study. In the most frequently visited places, participants remembered the location 72\% of the time, while only remembering the location 36\% of the time in least frequently visited places. This effect does not occur for trip memorability. In the most frequently visited places, trips are remembered 51\% of the time. Instead of seeing trip memorability decrease for least frequently visited places, this factor increases to 56\%.

@@

\subsubsection{Travel Companions}
Our findings show that trip remembrance is influenced by the presence of travel companions. As one might imagine, trips that were taken with other people were more memorable than trips taken alone. Participants were not able to remember the exact trip of 33\% of Rewinds taken when they were traveling alone, but they could remember the exact trip of 89\% of Rewinds that were taken with company.

\subsubsection{GPS Usage}
Our findings show that participants who are not regular GPS users remember locations and trips more often than participants who are regular GPS users (Figure~\ref{fig:gpsmemory}). When asked whether they remember the location or trip being shown in a photo or Rewind, participants who do not use GPS regularly gave a positive answer 58\% of the time for location remembrance and 56\% of the time for trip remembrance. Regular GPS users remembered locations 33\% of the time and trips 47\% of the time. The data shows that participants who are regular GPS users did not remember the locations they visited or trips they took as often as participants who are not regular GPS users. Since regular GPS users are frequently directed step-by-step on how to navigate their environment, they may not pay as much attention to their surroundings. This diminished locational awareness could be the reason why regular GPS users had a lower remembrance rate for locations and trips. From these percentages, we also see that regular GPS users remember trips more often than they remember locations, while there is little difference between the two for participants who do not use GPS regularly.

@@

\subsubsection{Frequent Travelers}
Our findings show that trip remembrance is influenced by the presence of travel companions. As one might imagine, trips that were taken with other people were more memorable than trips taken alone. Participants were not able to remember the exact trip of 33\% of Rewinds taken when they were traveling alone, but they could remember the exact trip of 89\% of Rewinds that were taken with company.

>>>>

			\multicolumn{1}{|c|}{}                               & \multicolumn{1}{c|}{}                        & \multicolumn{1}{c|}{mean}  & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}33.33\%}      & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}57.98\%}     & \multicolumn{1}{c|}{50.83\%}                              & \multicolumn{1}{c|}{47.46\%}                             & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}100.00\%}          & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}55.00\%}        & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}100.00\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}75.00\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}72.22\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}43.66\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}51.56\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}10.00\%} & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}72.34\%}      & \multicolumn{1}{c|}{\cellcolor[HTML]{ACDDAA}35.80\%}     \\ \cline{3-17} 

@@

% shall we change photo/rewind video rememberance to location/trip rememberance?
\subsection{Memorability}
Rewinds are meant to engage a person's memory. To find out more about what features make a digital memory memorable, we examined participant responses to whether or not they were able to remember the location shown in a photo or recall the trip portrayed by a Rewind. Photo memorability is determined by whether or not the participant remembered the location depicted in a photo. Because a Rewind contains many locations, it would be unfair to consider the memorability of Rewind based on whether participant remembers any locations in the Rewind. So, Rewind memorability is determined by whether or not the participant remembered and could distinguish the particular trip portrayed in a Rewind. In this section, we will look at differences in photo and Rewind memorability based on the variables of distance from the participant's home, visit frequency, presence of travel companions, regular GPS usage, and whether the participant considered themselves a frequent traveler.

% didn't do the stats for video location vs photo locations, not sure if it is necessary?
\subsubsection{Rewinds versus Photos}
When participants were shown a street side photo from their location history, they were asked whether they recognized the location in the photo. In 50\% of cases, participants said they did. When they were shown the Rewind associated with that photo, nearly every case had recognizable locations---97\%. This result is expected since Rewinds contain more location information than static photos, so if participants do not recognize some depicted locations, they still have many other chances to pick out locations they remember. Rewind memorability as calculated by whether participants could identify the particular trip they were on from watching the Rewind was 54\%. However, there were cases where the participant would only remember the photo after seeing its accompanying Rewind, which provided a context. More than one participant exclaimed after watching a Rewind, ``I recognize the photo now!''

\subsubsection{GPS Usage}
% Our findings show that GPS usage significantly affect the rememberance of locations $\chi^2(1,179) = 8.734$, p < 0.01, but does not impose a siginificant effect on trip rememberance. (GPS users remember locations 33% of the time whereas non-GPS users remember locations 58% of the time)
Our findings show that participants who are not regular GPS users remember locations and trips more often than participants who are regular GPS users (Figure~\ref{fig:gpsmemory}). When asked whether they remember the location or trip being shown in a photo or Rewind, participants who do not use GPS regularly gave a positive answer 58\% of the time for location remembrance and 56\% of the time for trip remembrance. Regular GPS users remembered locations 33\% of the time and trips 48\% of the time. The data shows that participants who are regular GPS users did not remember the locations they visited or trips they took as often as participants who are not regular GPS users. Since regular GPS users are frequently directed step-by-step on how to navigate their environment, they may not pay as much attention to their surroundings. This diminished locational awareness could be the reason why regular GPS users had a lower remembrance rate for locations and trips. From these percentages, we also see that regular GPS users remember trips more often than they remember locations, while there is little difference between the two for participants who do not use GPS regularly.

\subsubsection{Frequent Travelers}
% has no siginifant effects on location no trip rememberance
Our findings show that trip remembrance is influenced by the presence of travel companions. As one might imagine, trips that were taken with other people were more memorable than trips taken alone. Participants were not able to remember the exact trip of 33\% of Rewinds taken when they were traveling alone, but they could remember the exact trip of 89\% of Rewinds that were taken with company.

\subsubsection{Travel Companions}
% trip accompanies have siginificant effects on both location and trip rememberance. 
% location rememberance: travel alone -  100%, not alone - 55%, $\chi^2(1,94) = 31.812$, p < 0.01
% trip rememberance: travel alone - 62%, not alone - 90%, $\chi^2(1,133) = 9.016$, p < 0.01
Our findings show that trip remembrance is influenced by the presence of travel companions. As one might imagine, trips that were taken with other people were more memorable than trips taken alone. Participants were able to remember the exact trip of 67\% of Rewinds taken when they were traveling alone, but they could remember the exact trip of 90\% of Rewinds that were taken with company, with $\chi^2(1,133) = 9.016$, p < 0.01.

\subsubsection{Distance}
% only affect location rememberance
% 0 - 100%, 1 - 75%, 2 - 72%, 3 - 44%, 4 - 52%, 5 - 10%, $\chi^2(5,179) = 18.106$, p < 0.01
The distance away from the participant’s home had differing effects on photo memorability and Rewind memorability (Figure~\ref{fig:distancememory}). Distances ranged from 0 to 206,155 meters. For analysis, these distances were logarithmically scaled and normalized into six distance groups to mimic human categorization of distances into generalizations like ``close,'' ``far,'' or ``very far.'' When normalized: Group 0 corresponded to a range of 0--10 meters from the participant's home location; Group 1 corresponded to a range of 10--100 meters; Group 2 corresponded to a range of 100--1,000 meters; Group 3 corresponded to a range of 1,000--10,000 meters; Group 4 corresponded to a range of 10,000--100,000 meters and Group 5 corresponded to a range of 100,000--1,000,000 meters from the participant's home location.

Photo memorability and distance had a clear linear relationship. As the distance increased, photo memorability decreased, suggesting that locations further away from home were less likely to be remembered. Remembrance was 100\% for Distance Group 0, or home. One participant recognized a photo because her ``husband's car is in the driveway.'' Remembrance dropped to 75\% for Distance Group 1, then down to 10\% for Distance Group 5, which represented the locations furthest away from home. These percentages are statistically significant, with a p-value of 0.0028.

@@

\subsubsection{Visit Frequency}
% only affect location rememberance
% most frequently visited - 72%, least frequently visited - 36%, $\chi^2(1,128) = 14.458$, p < 0.01
Location memorability and trip memorability are affected differently by frequency of visits (Figure~\ref{fig:frequencymemory}). Location memorability is highest for places that are most frequently visited by the participant, while trip memorability is highest for places that are least frequently visited. The more frequently a place was visited, the more likely it was that the participant remembered the location when it was shown to them at the study. In the most frequently visited places, participants remembered the location 72\% of the time, while only remembering the location 36\% of the time in least frequently visited places, with a significance at $\chi^2(1,128) = 14.458$, p < 0.01. This effect does not occur for trip memorability. In the most frequently visited places, trips are remembered 51\% of the time. Instead of seeing trip memorability decrease for least frequently visited places, this factor increases to 56\%.

@@

@@
